Cannibalism among the supermassive black holes

### 2019-01-19 paper.tex
Added proper units to Black Hole Luminosity Section
Added Galaxy Gas X-Ray Luminosity section

\documentclass[english, apj]{emulateapj}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{rotating}
\usepackage{units}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{url}
\usepackage{babel}
\usepackage{float}
\usepackage{savesym}
\savesymbol{tablenum}
\usepackage{siunitx}
\restoresymbol{SIX}{tablenum}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage[all]{hypcap}

@@

The merger tree data from Lackner(+12) and Kulier(+15) provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4 > z > 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

@@

We also compared the stellar masses from Lackner(+12) to  observational and abundance matching data from \citet{2018AstL...44....8K}. Figure \ref{fig:stellar1} is a partial, approximate, replication of their Figure 11 (abundance matching line).  We rescaled our stellar masses at $z=0$ in a lognormal manner about the line in order to match the observed data.  Table \ref{table:gal_char} shows the original and rescaled stellar masses.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/stellar_to_halo_ratio.png}
\caption{Approximation of Figure 11 from \citet{2018AstL...44....8K}.  Purple dots shown is example of our rescaled stellar masses}
\label{fig:stellar1}
\end{center}
\end{figure}

\begin{table}[htbp]
\caption{Galaxy characteristics at z=0}
\begin{tabular} {|c | c | c| c|}
\hline
$M_{DM}$ & $M_{CBH}$ & $M_{*} (orig.)$ & $M_{*} (rescaled)$ \\
\hline
 \num{9.18e13} 	&	 \num{2.19e9} 	&	\num{1.75e13} 	&	\num{5.14e11}  \\
 \num{7.75e13} 	&	 \num{5.27e9}	&	\num{1.52e13} 	&	\num{4.00e11}  \\
 \num{2.85e13} 	&	 \num{1.46e9}	&	\num{5.58e12} 	&	\num{3.30e11}  \\
 \num{2.48e13} 	&	 \num{2.40e9}	&	\num{4.80e12} 	&	\num{2.25e11}  \\
 \num{2.29e13} 	&	 \num{1.22e9}	&	\num{4.46e12} 	&	\num{2.68e11}  \\
 \num{1.18e13} 	&	 \num{3.62e9}	&	\num{2.32e12} 	&	\num{2.83e11}  \\
 \num{1.03e13} 	&	 \num{9.85e9}	&	\num{1.96e12} 	&	\num{1.62e11}  \\
 \num{1.00e13} 	&	 \num{9.32e9}	&	\num{1.95e12} 	&	\num{2.49e11}  \\
 \num{8.92e12} 	&	 \num{5.60e9}	&	\num{1.74e12} 	&	\num{3.20e11}  \\
 \num{8.56e12} 	&	 \num{6.86e9}	&	\num{1.62e12} 	&	\num{1.83e11}  \\
 \num{6.94e12} 	&	 \num{2.02e9}	&	\num{1.72e12} 	&	\num{1.51e11}  \\
 \num{6.29e12} 	&	 \num{1.22e9}	&	\num{1.23e12} 	&	\num{1.87e11}  \\
 \num{1.84e12} 	&	 \num{2.21e9}	&	\num{5.52e11} 	&	\num{6.42e10}  \\
\hline
\end{tabular}

@@

\subsection{Simulation setup}
For each galaxy in the simulations, we only used black holes that would theoretically have merged with the center of the host galaxy within a dynamical friction time, $t_{fric}$, of less than 100 times a Hubble time, with $t_{fric}$ defined from \citet{2008gady.book.....B} as:
\begin{equation}\label{tfric}
    t_{fric} = \frac{19}{6}Gyr\left(\frac{R_e}{5kpc}\right)^2\frac{\sigma_{eff}}{200km\,s^{-1}}\frac{10^8\,M_{\odot}}{M_{bh}}
\end{equation}
The orbiting black holes were injected into their host galaxy at redshift z, at a distance from the galactic center of $R_{e}$ (Eqn. \ref{re}) following galaxy mergers.  Their initial velocity was arbitrarily chosen to be circular, $v_c = \sqrt{M(R_e)/R_e}$, with $v_x$, $v_y$, and $v_z$ randomly chosen.

\subsection{Black Hole Luminosity}
The x-ray temperature versus sigma relation from Table 2 of  \citet{2018ApJ...857...32B} was fitted according to:
\begin{equation}
    T_x = 10^{(2.50*log \sigma - 6.06)}
\end{equation}
Central gas density, ${\rho}_0$ was also fitted from Table 3 of \citet{2018ApJ...857...32B} according to:
\begin{equation}
    \rho_0 = 10^{(0.6log(\sigma) - 25)}
\end{equation}
The core radius of the gas model, ${r_c}$ is:
\begin{equation}
    r_c = 10^{(15.09log(\sigma) - 35.77)}
\end{equation}
Local gas density ${\rho}$ is obtained from the isothermal beta model:
\begin{equation}
    \rho = \frac{\rho_0}{(1+r^2/r_c^2)^{1.5}}
\end{equation}
Gas sound speed, $c_s$, was obtained from:
\begin{equation}
    c_s = \sqrt{1.67T_x/m_p},
\end{equation}
where $m_p$ is the proton mass.

The black hole mass accretion rate is approximated by:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{(c_s^2+v^2)^{3/2}},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

Finally, the black hole xray luminosity is calculated from:
\begin{equation}
    L_x = 0.01\dot{M}_{acc}c^2
\end{equation}

The Eddington limit is known to be:
\begin{equation}
    L_{edd} = \num{1.26e38} (M/M_{\odot}) erg/s
\end{equation}

>>>>

\documentclass[english, apj]{emulateapj}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amstext}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
%\usepackage[nobottomtitles]{titlesec}
\usepackage{array}
\usepackage{rotating}
\usepackage{units}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{url}
\usepackage{babel}
\usepackage{float}
\usepackage{savesym}
\savesymbol{tablenum}
\usepackage{siunitx}
\restoresymbol{SIX}{tablenum}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage[all]{hypcap}

@@

The merger tree data from Lackner(+12) and Kulier\-(+15) provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4 > z > 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

@@

We also compared the stellar masses from Lackner\-(+12) to  observational and abundance matching data from \citet{2018AstL...44....8K}. Figure \ref{fig:stellar1} is a partial, approximate, replication of their Figure 11 (abundance matching line).  We rescaled our stellar masses at $z=0$ in a lognormal manner about the line in order to match the observed data.  Table \ref{table:gal_char} shows the original and rescaled stellar masses.

\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/stellar_to_halo_ratio.png}
\caption{Approximation of Figure 11 from \citet{2018AstL...44....8K}.  Purple dots shown is example of our rescaled stellar masses}
\label{fig:stellar1}
\end{center}
\end{figure}

\begin{table}
\caption{Galaxy characteristics at z=0}
\begin{tabular} {|c | c | c| c|}
\hline
$M_{DM}$ & $M_{CBH}$ & $M_{*} (orig.)$ & $M_{*} (rescaled)$ \\
\hline
 \num{9.18e13} 	&	 \num{2.19e9} 	&	\num{1.75e13} 	&	\num{5.14e11}  \\
 \num{7.75e13} 	&	 \num{5.27e9}	&	\num{1.52e13} 	&	\num{4.00e11}  \\
 \num{2.85e13} 	&	 \num{1.46e9}	&	\num{5.58e12} 	&	\num{3.30e11}  \\
 \num{2.48e13} 	&	 \num{2.40e9}	&	\num{4.80e12} 	&	\num{2.25e11}  \\
 \num{2.29e13} 	&	 \num{1.22e9}	&	\num{4.46e12} 	&	\num{2.68e11}  \\
 \num{1.18e13} 	&	 \num{3.62e9}	&	\num{2.32e12} 	&	\num{2.83e11}  \\
 \num{1.03e13} 	&	 \num{9.85e9}	&	\num{1.96e12} 	&	\num{1.62e11}  \\
 \num{1.00e13} 	&	 \num{9.32e9}	&	\num{1.95e12} 	&	\num{2.49e11}  \\
 \num{8.92e12} 	&	 \num{5.60e9}	&	\num{1.74e12} 	&	\num{3.20e11}  \\
 \num{8.56e12} 	&	 \num{6.86e9}	&	\num{1.62e12} 	&	\num{1.83e11}  \\
 \num{6.94e12} 	&	 \num{2.02e9}	&	\num{1.72e12} 	&	\num{1.51e11}  \\
 \num{6.29e12} 	&	 \num{1.22e9}	&	\num{1.23e12} 	&	\num{1.87e11}  \\
 \num{1.84e12} 	&	 \num{2.21e9}	&	\num{5.52e11} 	&	\num{6.42e10}  \\
\hline
\end{tabular}

@@

\subsection{Simulation setup}
For each galaxy in the simulations, we only used black holes that would theoretically have merged with the center of the host galaxy within a dynamical friction time, $t_{fric}$, of less than 100 times a Hubble time, with $t_{fric}$ defined from \citet{2008gady.book.....B} as:
\begin{equation}\label{tfric}
    t_{fric} = \frac{19}{6}Gyr\left(\frac{R_e}{5kpc}\right)^2\frac{\sigma_{eff}}{200km\,s^{-1}}\frac{10^8\,M_{\odot}}{M_{bh}} \text{  s}
\end{equation}
The orbiting black holes were injected into their host galaxy at redshift z, at a distance from the galactic center of $R_{e}$ (Eqn. \ref{re}) following galaxy mergers.  Their initial velocity was arbitrarily chosen to be circular, $v_c = \sqrt{M(R_e)/R_e}$, with $v_x$, $v_y$, and $v_z$ randomly chosen.

\subsection{Black Hole Luminosity}
The x-ray temperature versus sigma relation from Table 2 of  \citet{2018ApJ...857...32B} was fitted according to:
\begin{equation}
    T_x = 10^{2.50*log(\sigma) - 6.06} \text{  kev}
\end{equation}
Central gas density, ${\rho}_0$ was also fitted from Table 3 of \citet{2018ApJ...857...32B} according to:
\begin{equation}
    \rho_0 = 10^{0.6*log(\sigma) - 25} \text{  g cm}^{-3}
\end{equation}
The core radius of the gas model, ${r_c}$ is:
\begin{equation}
    r_c = 10^{2.1*log(\sigma) - 5.3} \text{  cm}
\end{equation}
Local gas density ${\rho}$ is obtained from the isothermal beta model:
\begin{equation} \label{beta_model}
    \rho = \frac{\rho_0}{(1+r^2/r_c^2)^{1.5}} \text{  g cm}^{-3}
\end{equation}
Gas sound speed, $c_s$, was obtained from:
\begin{equation}
    c_s = \sqrt{1.67T_x/m_p} \text{  km s}^{-1},
\end{equation}
where $m_p$ is the proton mass.

The black hole mass accretion rate is approximated by:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{(c_s^2+v^2)^{3/2}} \text{  M}_\odot \text{ Myr}^{-1},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

Finally, the black hole xray luminosity is calculated from:
\begin{equation}
    L_x = 0.01\dot{M}_{acc}c^2 \text{  erg s}^{-1}
\end{equation}

The Eddington limit is known to be:
\begin{equation}
    L_{edd} = \num{1.26e38} (M/M_{\odot}) \text{  erg s}^{-1}
\end{equation}

\subsection{Galaxy Gas X-Ray Luminosity}
We compute the x-ray luminosity of the hot gas in the galaxies, adapting Equation 26 from \citet{2012ApJ...754..125C}. In lieu of summing over discrete particles, we use the beta model of the gas as defined in our equation \ref{beta_model} and integrate:
\begin{equation}
\begin{aligned}
    L_x ={} \frac{\num{1.2e-24}}{(\mu m_p)^2}\left(\frac{kT_x}{1keV}\right)^{1/2}\rho_0^{2}\;4{\pi} \\
    \int_{0}^{\infty}\frac{r^2}{(1+r^2/r_c^2)^{3}}\;dr\text{  erg s}^{-1}
\end{aligned}
\end{equation}
The integral evaluates to ${\pi}r_c^3/16$.  $m_p$ is the mass of a proton and $\mu$ is the mean molecular weight.  To calculate $\mu$, we use metallicity $Z=2Z_{\odot}=0.0268$, mass fraction of Helium $Y=Y_{\odot}=0.2485$, and mass fraction of Hydrogen $X=0.7247$.  This results in $\mu=0.606$.
